Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-02-22 alloscore_manuscript/alloscore-application-shorter.Rnw
some edits to manuscript

The \textbf{allocation score} of the forecast $F$ is the avoidable unmet need that results from using the allocation
that minimizes the expected unmet need according to that forecast. By ``avoidable unmet need'', we mean that the
allocation score does not include the amount of unmet need that was inevitable simply because the amount of available
resources $K$ was less than the need for resources. Rather, the allocation score measures the unmet need that could have
been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so
that nothing is wasted in one location while it could be put to use in another. The best and lowest possible allocation 
score is 0.  A positive allocation score indicates some way in which an oracle could have used perfect foreknowledge to
meet more of the need than did the allocation suggested by $F$.

@@

Next suppose that we observe resource needs of 1 and 10 in locations 1 and 2, respectively. Based on these observed
needs, we evaluate the allocation suggested by the forecast by calculating the amount of unmet need that resulted from
that allocation over and above what was unavoidable given the resource constraint. With $K = 5$ available resource
units, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6 units of need
unmet in location 2. But these 6 units of unmet need are unaviodable under the constraint $K=5$: allocating 0 units of
resources to location 1 and 5 to location 2, for example, still results in a total unmet need of 6 across both
locations. This gives an allocation score of 0 for the forecast when $K = 5$. On the other hand, when $K = 10$, the
forecast's allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need unmet in
location 1. In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from going unmet,
for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to location 2. So for
$K = 10$, the forecast gets an allocaton score of 1 (= 2 realized $-$ 1 unavoidable) in units of avoidable unmet need.

@@

Decision theory is concerned with understanding how an \emph{action} $x$ both is and ought to be selected from a set of
possible actions $\mathcal{X}$ --- that is, how a decision is and ought to be made --- based on the anticipated
consequences of $x$ under future states. We represent such future states as a random variable $Y$ and refer to a
realization $y$ of $Y$ as an \emph{outcome}. In this paper $x$ will be related to the level of investment in a measure
designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with
$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ will correspond to the number of
individuals who eventually become sick and would benefit from the mitigation measure. A basic strategy in decision
theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
\emph{loss}, measuring the success of $x$, which in our examples will be the extent to which $x$ meets the realized need
$y$.  One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
$Y$. The loss of the action $x$ chosen according to $F$ can, in turn, be interpreted as the value of $F$ as an input to
the decision making process under the realized outcome $y$ of $Y$.

We arrive through this discussion at a three-step procedure for developing scoring rules for probabilistic forecasts:
\begin{enumerate}
  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$
    eventually occurs.
  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under
    the distribution $F$.
  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: 
    $S(F, y) = s(x^F, y)$.
\end{enumerate}
This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} --- a terminology we find to
be convenient but which is not, to our knowledge, standard.  In appendix \ref{sec:proper} we review the
basic fact from the scoring rule literature (e.g., \cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring
rules are proper by construction.

@@

In our setting an action is a vector $x = (x_1, \ldots, x_N)$ specifying the amount of resources allocated to each of
$N$ locations. We require that $0 \leq x$, i.e., that each $x_i$ is non-negative, and that the total allocation across
all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$. The set $\mathcal{X}$ consists of
all possible allocations that satisfy these constraints. The eventually realized resource need in each location is
denoted by $y = (y_1, \ldots, y_N)$. The future outcome that will determine the loss caused by choosing the allocation
$x$ is the vector of resource needs $y = (y_1, \ldots, y_N)$ for each location which is a realization (unknown at the
time of decision making) of the random vector $Y = (Y_1, \ldots, Y_N)$. A forecast of the random need $Y$ for each
location is written as $F = (F_1, \ldots, F_N)$. We assume need is non-negative and the forecasts reflect that, i.e. the
support of each $F_i$ is a subset of $\mathbb{R}^+$. Finally, we assume that any unmet unit of need in any location
incurs a fixed loss of $L>0$, so that if the selected resource level $x_i$ in location $i$ is less than the realized
need $y_i$, a loss of $L \cdot (y_i - x_i)$ results. A variety of extensions to this setup are possible; for example, we
might account for storage costs for resources that go unused, allow for a different loss per unit of unmet need in each
location, or account for resource transportation costs. In this work, we choose to keep the loss function relatively
simple to focus on the core ideas.

It is helpful to clearly distinguish between the time $t_d$ when a \emph{d}ecision is made about a public health
resource allocation and the time $t_r$ when \emph{r}esource needs that might be addressed by that allocation occur. Our
setup assumes that $t_d < t_r$ and that the resource in question can only meet the need realized at $t_r$, not prevent
it. That is, we stipulate that the random vector $Y$ of need is independent of the chosen allocation $x$. In the
context of infectious disease, this means that we do not consider resources, such as vaccines, that are intended to
reduce the number of people who will become sick at some point in the future. Instead, our setup addresses resources
like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are
already sick. Additionally, our setup addresses decision-making that is related to resource needs only at the time
$t_r$; we do not consider sequences of multiple decisions that are made over time or account for the impact of decisions
on resource needs at any time other than $t_r$. We outline some opportunities to extend our work to more complex
decision making settings in the discussion.

@@

<<thermometer-plot, fig.height=7, fig.width=7.5, fig.cap="Probabilistic forecasts for new COVID-19 hospital admissions on January 3, 2022 and their suggested resource allocations for the states with the ten highest (eventually observed) hospitalization counts. For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts. The side-by-side shaded regions show the median (solid red horizontal line) and 50\\%, 80\\% and 95\\% prediction intervals for the two selected models. The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data). The vertical bars with purple, blue and red shading show the allocations. The purple bar goes from zero to the amount of need that was met by the allocation from that model to a specific location. A red bar indicates need that exceeded the resource allocation for a location, and a light blue bar shows the amount by which the resource allocation suggested by that model exceeded the need for that location.">>=
source("R/plot_functions.R")

@@

A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the
\texttt{JHUAPL-Bucky} and \texttt{CU-select} models shows that while the \texttt{JHUAPL-Bucky} forecast distributions
were more centered on the eventual observations in many states, the suggested allocations of the \texttt{JHUAPL-Bucky}
model were less efficient than those of the \texttt{CU-select} model. Thus \texttt{JHUAPL-Bucky} achieved a worse
allocation score than \texttt{CU-select} (on this particular week) by allocating excess resources to several states,
such as Pennsylvania and Michigan, rather than to states such as Florida and California where these resources would have
reduced unmet need.  These allocation errors of \texttt{JHUAPL-Bucky} resulted from forecasts failing to consistently
capture the relative resource needs across different states. The \texttt{CU-select} model made some similar errors —
most prominently, over-allocating resources to Ohio — but overall, it more successfully forecasted the resource demands
across different locations in relative terms.

@@

<<multi-loc-ranks, fig.width=8, fig.cap="Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (\\texttt{JHUAPL-Bucky} and \\texttt{CU-select}). Panel A shows the observed resource need, in this case the observed number of hospitalizations, for each state, along with the hypothetical number of resources allocated to the given location based on the forecasts from each model. The number of available resources was fixed at 15,000 and forecasts from each model were used to determine an optimal allocation strategy before the resource need was known. For most locations the resource need exceeded the resources allocated, indicated by the `observed resource need' bar being larger than the `allocation' bar. Panel B shows the breakdown of the weighted interval score (WIS) into components of underprediction, overprediction and dispersion. Larger values of WIS indicate more error, and the full WIS score for each location can be decomposed into the three components shown here.">>=
## date points to target_end_date of 2022-01-10 to highlight
ref_date_for_plot <- "2021-12-20"
models_to_plot <- c("JHUAPL-Bucky", "CU-select", "per-capita")

@@

<<metrics-over-time, fig.height=8, fig.width=8, fig.cap="Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the dates for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the hypothetical resource constraint $K$. Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states. The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon). AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). In Panel B, the \\texttt{per-capita} allocation is drawn in a heavier solid line. Panel C shows the MWIS metric across weeks, averaged across all states. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.">>=
p2 <- ggplot(alloscore_data_summaries,
             aes(x=target_end_date, y=alloscore, color=model, shape=model, linewidth = model)) +
  geom_point() + geom_line()+
  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
  scale_color_manual(values = model_palette) +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
  scale_linewidth_manual(values = model_linewidths) +
  ylab("Allocation score") +
  theme(legend.pos="none") +
  xlab(NULL)

@@

Our example application of the AS refers to a hypothetical and unspecified resource need that is assumed to pair with
our forecasted outcome (new hospital admissions) in a manner that might appear overly simplistic and artificially
direct.  We are optimistic, however, that by augmenting this scheme with a probabilistic measurement model for resource
need and more detailed domain knowledge, it could be meaningfully aligned with a more concrete decision problem such as
the allocation of ventilators during a respiratory viral pandemic.  This problem was explored from a stochastic
optimization perspective (not from a forecast evaluation perspective) in \cite{huang_stockpiling_2017} and provided an
initial motivation for this project. We see it as promising focus for further development of the AS framework since it
(a) has been identified as a prime manifestation of scarcity during respiratory outbreaks (b) tansparently involves
logistical considerations depending on geographical and population scales, and (c) would likely imply a central role for
hospitalization data. But again, additional modeling and data synthesis would be required since not everyone who is
hospitalized needs a ventilator, and new hospital admissions will not fully describe for a need that corresponds to a
prevalence of severe cases. Other possible real-world examples include the allocation of a limited stockpile of
vaccinations \citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests
\citep{du_optimal_2022,pasco_covid-19_2023}.

@@

In conclusion, we argue that the way modelers and policymakers view and evaluate forecasts should more explicitly depend on
the specific decision-making context. Defaulting to standard forecast evaluation metrics can mask the utility (or
disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those
for which they are able to offer useful guidance. New collaborative work between public health officials and modeling teams is
needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or
simulation exercises that could be used to inform further development of new or alternative scoring metrics. We see this
work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling
applied epidemiological forecasting with public health decision making.

@@

In order for $x^{\star} \in \mathbb{R}^N_{+}$ to solve the AP it must be true that reallocating $\delta > 0$ units of
the resource from location $i$ to location $j$ will lead to a net increase in expected shortage --- in other words, the
reallocation increases the expected shortage in location $i$ is at least as much as it decreases the expected shortage
in location $j$:
\begin{align*}
&\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}]
\text{ (increase in $i$) } \nonumber \\
&\qquad \geq
\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]
\text{ (decrease in $j$) }.
\end{align*}

@@

In section \ref{sec:application} of the manuscript, we used the allocation score to evaluate forecasts of COVID-19
hospitalizations that have been submitted to the US COVID-19 Forecast Hub. These forecasts are submitted to the Hub
using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01,
0.025, 0.05, 0.1, 0.15, \ldots, \allowbreak 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the
endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2,
\allowbreak 0.3, 0.4, 0.5, 0.6, \allowbreak 0.7, 0.8, 0.9$. For a given week and target date, we use $q_{i,k}$ to denote
the submitted quantiles for location $i$ and probability level $\tau_k \in \mathcal{T}$, $k = 1, \ldots, 23$.

@@

To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need
a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided
quantiles. We have developed functionality for this purpose in the \verb`distfromq` package for R.\citep{ray-distfromq}
This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
\begin{enumerate}
  \item Identify a discrete component of the distribution consisting of zero or more point masses, and create an
    adjusted set of predictive quantiles for the continuous part of the distribution by subtracting the point mass
    probabilities and rescaling.
  \item For the continuous part of the distribution, different approaches are used on the interior and exterior of the
    provided quantiles:
  \begin{enumerate}
    \item On the interior, a monotonic cubic spline interpolates the adjusted quantiles representing the continuous part
      of the distribution.
    \item A location-scale parametric family is used to extrapolate beyond the provided quantiles. The location and
      scale parameters are estimated separately for the lower and upper tails so as to obtain a tail distribution that
      matches the two most extreme quantiles in each tail. In this work, we use normal distributions for the tails.
  \end{enumerate}
\end{enumerate}
The resulting distributional estimate exactly matches all of the predictive quantiles provided by the forecaster. We use
the cumulative distribution function resulting from this procedure as an input to the allocation score algorithm.

>>>>

The \textbf{allocation score} of the forecast $F$ is the avoidable unmet need that results from using the allocation
that minimizes the expected unmet need according to that forecast. By ``avoidable unmet need'', we mean that the
allocation score does not include the amount of unmet need that was inevitable simply because the amount of available
resources $K$ was less than the need for resources. Rather, the allocation score measures the unmet need that could have
been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so
that nothing is wasted in one location while it could be put to use in another. The best and lowest possible allocation 
score is 0.  A positive allocation score indicates that an oracle with perfect foreknowledge could have selected an allocation 
that met more need than the allocation suggested by $F$.

@@

Next suppose that we observe resource needs of 1 and 10 in locations 1 and 2, respectively. Based on these observed
needs, we evaluate the allocation suggested by the forecast by calculating the amount of unmet need that resulted from
that allocation over and above what was unavoidable given the resource constraint. With $K = 5$ available resource
units, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6 units of need
unmet in location 2. But these 6 units of unmet need are unavoidable under the constraint $K=5$: allocating 0 units of
resources to location 1 and 5 to location 2, for example, still results in a total unmet need of 6 across both
locations. This gives an allocation score of 0 for the forecast when $K = 5$. On the other hand, when $K = 10$, the
forecast's allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need unmet in
location 1. In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from going unmet,
for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to location 2. So for
$K = 10$, the forecast gets an allocaton score of 1 (= 2 realized $-$ 1 unavoidable) in units of avoidable unmet need.

@@

Decision theory is concerned with how to decide on an \emph{action} $x$ to select from a set of
possible actions $\mathcal{X}$ in the face of uncertainty about the future.
We use the random variable $Y$ to represent future states and refer to a
realization $y$ of $Y$ as an \emph{outcome}. For example, $x$ might correspond to the level of investment in a measure that is
designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with
$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ might correspond to the number of
individuals who eventually become sick and would benefit from the mitigation measure. A foundational strategy in decision
theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numeric value, or
\emph{loss}, measuring the (lack of) success of $x$. In our example, the loss measures the extent to which $x$ meets the realized need
$y$, with greater unmet need incurring a larger loss.
One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
$Y$.
This can help inform the selection of the action to take; for example, the decision maker might choose an action with the lowest expected loss according to the forecast.
The loss of the action $x$ chosen according to $F$ can, in turn, be interpreted as the value of $F$ as an input to
the decision making process under the realized outcome $y$ of $Y$.

Through this discussion, we arrive at a three-step procedure for developing scoring rules for probabilistic forecasts:
\begin{enumerate}
  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$
    eventually occurs.
  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under
    the distribution $F$.
  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: 
    $S(F, y) = s(x^F, y)$.
\end{enumerate}
This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} (noting that to our knowledge, this terminology is not standard).  In appendix \ref{sec:proper} we review the
result from the scoring rule literature (e.g., \cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring
rules are proper by construction.

@@

In our setting an action is a vector $x = (x_1, \ldots, x_N)$ specifying the amount of resources allocated to each of
$N$ locations. We require that $0 \leq x$, i.e., that each $x_i$ is non-negative, and that the total allocation across
all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$. The set $\mathcal{X}$ consists of
all possible allocations that satisfy these constraints. The eventually realized resource need in each location is
denoted by $y = (y_1, \ldots, y_N)$; this is a realization (unknown at the
time of decision making) of the random vector $Y = (Y_1, \ldots, Y_N)$. A forecast of the random need $Y$ for each
location is written as $F = (F_1, \ldots, F_N)$. We assume need is non-negative and the forecasts reflect that, i.e. the
support of each $F_i$ is a subset of $\mathbb{R}^+$. Finally, we assume that each unmet unit of need in any location
incurs a fixed loss of $L>0$, so that if the selected resource level $x_i$ in location $i$ is less than the realized
need $y_i$, a loss of $L \cdot (y_i - x_i)$ results. A variety of extensions to this setup are possible; for example, we
might account for storage costs for resources that go unused, allow for a different loss per unit of unmet need in each
location, or account for resource transportation costs. In this work, we choose to keep the loss function relatively
simple to focus on the core ideas.

It is helpful to clearly distinguish between the time $t_d$ when a \emph{d}ecision is made about a public health
resource allocation and the time $t_r$ when \emph{r}esource needs that might be addressed by that allocation occur. Our
setup assumes that $t_d < t_r$ and that the resource in question can only meet the need realized at $t_r$, not prevent
it. That is, we stipulate that the choice of allocation $x$ has no effect on the distribution of the random vector $Y$. In the
context of infectious disease, this means that we do not consider resources, such as vaccines, that are intended to
reduce the number of people who will become sick at some point in the future. Instead, our setup addresses resources
like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are
already sick. Additionally, our setup addresses decision-making that is related to resource needs only at the time
$t_r$; we do not consider sequences of multiple decisions that are made over time or account for the impact of decisions
on resource needs at any time other than $t_r$. We outline some opportunities to extend our work to more complex
decision making settings in the discussion.

@@

<<thermometer-plot, fig.height=7, fig.width=7.5, fig.cap="Probabilistic forecasts for new COVID-19 hospital admissions on January 3, 2022 and their suggested resource allocations for the states with the ten highest (eventually observed) hospitalization counts. For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts. The side-by-side shaded regions show the median (grey horizontal line) and 50\\%, 80\\% and 95\\% prediction intervals for the two selected models. The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data). The vertical bars with purple, blue and red shading show the allocations. The purple bar goes from zero to the amount of need that was met by the allocation from that model to a specific location. A red bar indicates need that exceeded the resource allocation for a location, and a light blue bar shows the amount by which the resource allocation suggested by that model exceeded the need for that location.">>=
source("R/plot_functions.R")

@@

A state-wise comparison of the
\texttt{JHUAPL-Bucky} and \texttt{CU-select} models shows that while the \texttt{JHUAPL-Bucky} forecast distributions
were more centered on the eventual observations in many states, the suggested allocations of the \texttt{JHUAPL-Bucky}
model were less efficient than those of the \texttt{CU-select} model (Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A). Thus \texttt{JHUAPL-Bucky} achieved a worse
allocation score than \texttt{CU-select} (on this particular week) by allocating excess resources to several states,
such as Pennsylvania and Michigan, rather than to states such as Florida and California where these resources would have
reduced unmet need.  These allocation errors resulted from forecasts failing to consistently
capture the relative resource needs across different states. The \texttt{CU-select} model made some similar errors —
most prominently, over-allocating resources to Ohio — but overall, it more successfully forecasted the resource demands
across different locations in relative terms.

@@

<<multi-loc-ranks, fig.width=8, fig.cap="Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (\\texttt{JHUAPL-Bucky} and \\texttt{CU-select}). Panel A shows the observed resource need, in this case the observed number of hospitalizations, for each state, along with the hypothetical number of resources allocated to the given location based on the forecasts from each model. The number of available resources was fixed at 15,000 and forecasts from each model were used to determine an optimal allocation strategy before the resource need was known. For most locations the resource need exceeded the resources allocated, indicated by some amount of `unmet resource need' above the `need met by allocated resources'. Panel B shows the breakdown of the weighted interval score (WIS) into components of underprediction, overprediction and dispersion. Larger values of WIS indicate more error, and the full WIS score for each location can be decomposed into the three components shown here.">>=
## date points to target_end_date of 2022-01-10 to highlight
ref_date_for_plot <- "2021-12-20"
models_to_plot <- c("JHUAPL-Bucky", "CU-select", "per-capita")

@@

<<metrics-over-time, fig.height=8, fig.width=8, fig.cap="Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the dates for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the hypothetical resource constraint $K$. Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states. The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (i.e., the date the forecast was made plus the forecast horizon). AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). In Panel B, the \\texttt{per-capita} allocation is drawn in a heavier solid line. Panel C shows the MWIS metric across weeks, averaged across all states. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.">>=
p2 <- ggplot(alloscore_data_summaries,
             aes(x=target_end_date, y=alloscore, color=model, shape=model, linewidth = model)) +
  geom_point() + geom_line()+
  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
  scale_color_manual(values = model_palette) +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
  scale_linewidth_manual(values = model_linewidths) +
  ylab("Allocation score") +
  theme(legend.pos="none") +
  xlab(NULL)

@@

Our example application of the AS refers to a hypothetical and unspecified resource need that is assumed to pair with
our forecasted outcome (new hospital admissions) in a manner that might appear overly simplistic and artificially
direct.  We are optimistic, however, that by augmenting this scheme with a probabilistic measurement model for resource
need and more detailed domain knowledge, it could be meaningfully aligned with a more concrete decision problem such as
the allocation of ventilators during a respiratory viral pandemic.  This problem was explored from a stochastic
optimization perspective (not from a forecast evaluation perspective) in \cite{huang_stockpiling_2017} and provided an
initial motivation for this project. We see it as promising focus for further development of the AS framework since it
(a) has been identified as a setting in which allocation decisions are made during respiratory disease outbreaks, (b) involves
logistical considerations depending on geographical and population scales, and (c) would likely imply a central role for
hospitalization data. But again, additional modeling and data synthesis would be required since not everyone who is
hospitalized needs a ventilator. Other possible real-world examples include the allocation of a limited stockpile of
vaccinations \citep{araz_geographic_2012,persad_fair_2023} or diagnostic tests
\citep{du_optimal_2022,pasco_covid-19_2023}.

@@

In conclusion, we argue that when possible, the way modelers and policymakers view and evaluate forecasts should more explicitly depend on
the specific decision-making context. Defaulting to standard forecast evaluation metrics can mask the utility (or
disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those
for which they are able to offer useful guidance. New collaborative work between public health officials and modeling teams is
needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or
simulation exercises that could be used to inform further development of new or alternative scoring metrics. We see this
work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling
applied epidemiological forecasting with public health decision making.

@@

In order for $x^{\star} \in \mathbb{R}^N_{+}$ to solve the AP it must be true that reallocating $\delta > 0$ units of
the resource from location $i$ to location $j$ will lead to a net increase in expected shortage --- in other words, the
reallocation increases the expected shortage in location $i$ at least as much as it decreases the expected shortage
in location $j$:
\begin{align*}
&\mathbb{E}_{F_i}[(Y_i - (x^{\star}_i - \delta))_{+}] - \mathbb{E}_{F_i}[(Y_i - x^{\star}_i)_{+}]
\text{ (increase in $i$) } \nonumber \\
&\qquad \geq
\mathbb{E}_{F_j}[(Y_j - x^{\star}_j)_{+}] - \mathbb{E}_{F_j}[(Y_j - (x^{\star}_j + \delta))_{+}]
\text{ (decrease in $j$) }.
\end{align*}

@@

In section \ref{sec:application}, we used the allocation score to evaluate forecasts of COVID-19
hospitalizations that have been submitted to the US COVID-19 Forecast Hub. These forecasts are submitted to the Hub
using a set of 23 quantiles of the forecast distribution at the 23 probability levels in the set $\mathcal{T} = \{0.01,
0.025, 0.05, 0.1, 0.15, \ldots, \allowbreak 0.9, 0.95, 0.975, 0.99\}$, which specify a predictive median and the
endpoints of central $(1 - \alpha) \times 100\%$ prediction intervals at levels $\alpha = 0.02, 0.05, 0.1, 0.2,
\allowbreak 0.3, 0.4, 0.5, 0.6, \allowbreak 0.7, 0.8, 0.9$. For a given week and target date, we use $q_{i,k}$ to denote
the submitted quantiles for location $i$ and probability level $\tau_k \in \mathcal{T}$, $k = 1, \ldots, 23$.

@@

To address this situation and support the numerical allocation algorithm outlined in section \ref{sec:numeric}, we need
a mechanism to approximate the full cumulative distribution functions $F_i$, $i = 1, \ldots, N$ based on the provided
quantiles. We have developed functionality for this purpose in the \verb`distfromq` package for R \citep{ray-distfromq}.
This functionality represents a distribution as a mixture of discrete and continuous parts, and it works in two steps:
\begin{enumerate}
  \item Identify a discrete component of the distribution consisting of zero or more point masses, and create an
    adjusted set of predictive quantiles for the continuous part of the distribution by subtracting the point mass
    probabilities and rescaling.
  \item For the continuous part of the distribution, different approaches are used on the interior and exterior of the
    provided quantiles:
  \begin{enumerate}
    \item On the interior, a monotonic cubic spline interpolates the adjusted quantiles representing the continuous part
      of the distribution.
    \item A location-scale parametric family is used to extrapolate beyond the provided quantiles. The location and
      scale parameters are estimated separately for the lower and upper tails so as to obtain a tail distribution that
      matches the two most extreme quantiles in each tail. In this work, we use normal distributions for the tails.
  \end{enumerate}
\end{enumerate}
The resulting distributional estimate exactly matches all of the predictive quantiles provided by the forecaster. We use
the cumulative distribution function resulting from this procedure as an input to the allocation score algorithm.
